\section{Axioms}

This section states five axioms that characterize deterministic execution and observational separation.

\paragraph{A1 Atomicity of $\Delta$.}
Execution is composed of atomic events. No event can be partially applied or split into smaller execution units for the purposes of ordering or replay.

\paragraph{A2 Append-only ordered stream $V$.}
The execution stream $V$ is append-only and totally ordered by the index $t(e)$. Events are immutable once appended.

\paragraph{A3 Order is independent of wall clock time.}
The order induced by $t$ is defined by stream position only. Wall clock time is observational and does not affect order.

\paragraph{A4 Observations are non-derivational.}
Observations may be recorded and may vary arbitrarily. The canonical projection $\pi(V; I_{\text{exec}})$ must not depend on observations. Changes in observations alone cannot change $\pi(V; I_{\text{exec}})$.

\paragraph{A5 Structural determinism of projections.}
For fixed authoritative execution inputs $I_{\text{exec}}$, the canonical projection $\pi(V; I_{\text{exec}})$ is deterministic and invariant under execution replay. Authoritative execution inputs are those required to identify and order events in $V$ and to evaluate $\pi$; observational artifacts are excluded.

\paragraph{Reference Implementation.}
We provide a minimal reference implementation of the Deterministic Boundary
Layer (DBL) as a runtime system.\footnote{Implementation available at
\url{https://github.com/lukaspfisterch/dbl-gateway}.}

The system realizes the event chain
\textit{INTENT} $\rightarrow$ \textit{DECISION} $\rightarrow$ \textit{PROOF}
$\rightarrow$ \textit{EXECUTION}. DECISION events are recorded in an
append-only log prior to execution and constitute the normative layer.
Execution outputs are treated as non-normative observations and do not affect
governance decisions.

The implementation is organized to preserve the core invariants of the model:
(i) append-only event persistence,
(ii) strict ordering of DECISION before EXECUTION,
(iii) governance input purity, where decisions depend only on authoritative
inputs and policy configuration, and
(iv) deterministic replay of DECISION events under fixed inputs and policy
configuration.

The reference implementation provides a concrete instantiation of the model.
Empirical results on replay equivalence and policy variation are reported in
Section~\ref{sec:evaluation}.
